Phenomenological Robotics

### 2011-12-09 draft.tex
Wrote on Merleau-Ponty and relation to E. Gibson. Added dreyfus2002 and gibson2000 to bibliography.

>>>>

% 2. Relation between phenomenology and alternative psychology theories.

\section{Phenomenology and psychology theories}

% 2.B. Merleau-Ponty skills (refined situations) vs. Eleanor Gibson.
% (Source: Thesis journal, pages 140 to 149, 29/03/2011).
% (Source: \cite{gibson2000}.)

Dreyfus~\cite{dreyfus2002} shows that learning and skillful action, deemed as
basic forms of intelligent behavior, can be explained without recurring to
mental representations. To accomplish this, he explains the two central concepts
in Merleau-Ponty's \emph{Phenomenology of Perception}:

\begin{itemize}
\item \textbf{intentional arc:} is the strong connection between an agent and
its environment. In that sense, the skills learned by an agent are not kept as
representations in the mind, but as dispositions to respond to solicitations of
situations in the environment;
\item \textbf{maximal grip:} is the tendency of an agent to respond to these
solicitations in a way that brings the situation status closer to what it
considers optimal.
\end{itemize}

Gibson~\cite{gibson2000} propose that the process of learning is one of
discrimination rather than association, that is, through the refinement of the
sensorimotor knowledge acquired through the selection of patterns in our
interaction with the world. According to her, this selection process happens
based on two criteria:

% (\cite{gibson2000}, pages 154-156).

\begin{itemize}
\item \textbf{affordance fit:} what increases the fit between the organism and
its environment, according to its capacities and goals;
\item \textbf{economy:} what reduces the need for action and perceptual
information to accomplish a given task. -- optimization
\end{itemize}

These criteria can be seen as directly related to Merleau-Ponty's
\emph{intentional arc} and \emph{maximal grip}, respectively.
